---
title: "Workflow and Formatting"
author: "STA199 TA team"
format: html
editor: visual
---

### Rubrics for workflow and formatting

Breakdown of overall workflow and formatting grading listed below:

-   1 pts: 1 point given if all *code chunks are appropriately labelled*
    -   take off 0.5 pt if anything less
    -   take off 1 pt if none of the code chunks are appropriately labelled
-   3 pts: Consistent code style.
-   1 point given if narrative and text does not exceed the 80 character limit.
-   1 point given for consistent use of operators (i.e. %\>% or \|\> and \<- or = but not both)
-   1 point given for consistent and appropriate use of spacing (i.e. pipes and layers followed by newline and binary operators like + and - are surrounded by spaces)
-   1 pt: author name updated to the student's name at the YAML (at the top) of the document.

### Common Mistakes

#### 1. meaningless labels

```{r chunk-0}
library(tidyverse)
```

```{r chunk-1}
iris |>
  glimpse()
```

Some good examples

```{r iris-dataset}
iris |>
  glimpse()
```

```{r library}
iris |>
  glimpse()
```

If you have multiple models or plots for the homework/labs, you can number them.

```{r lm-1}
iris_sum <- iris |>
  as_tibble() |>
  lm(formula = Petal.Width ~ Sepal.Width) |>
  summary()
```

```{r boxplot-1}
iris_boxplot <- iris |>
  as_tibble() |>
  ggplot(aes(y = Petal.Width, x = Species)) + geom_boxplot()
```

#### 2. Inconsistent use of operator

Bad examples: use both %\>% and \|\>, both \<- and = in the homework/lab.

```{r bad-op-1}
iris_boxplot <- iris |>
  as_tibble() %>%
  ggplot(aes(y = Petal.Width, x = Species)) + 
  geom_boxplot()
```

```{r bad-op-2}
iris_tbl <- iris |>
  as_tibble()
iris_boxplot = ggplot(iris_tbl, aes(y = Petal.Width, x = Species)) + 
  geom_boxplot()
```

Good examples:

```{r good-op-1}
iris_boxplot <- iris |>
  as_tibble() |>
  ggplot(aes(y = Petal.Width, x = Species)) + 
  geom_boxplot()
```

```{r good-op-2}
iris_tbl <- iris |>
  as_tibble()
iris_boxplot <- ggplot(iris_tbl, aes(y = Petal.Width, x = Species)) + 
  geom_boxplot()
```

#### 3. Narrative and text exceeding character limit

Bad examples: text overflow. Only part of the text will appear in both coding part and plot.

```{r bad-txt-1}
iris |>
  as_tibble() |>
  ggplot(aes(y = Petal.Width, x = Species)) + 
  geom_boxplot() +
  labs(title = "The title is super long and exceeds the word limits. Please do not take my points off.")
```

Possible solutions: have a new line inside your code or have subtitle. It is always a good idea to do proofreading on your knitted pdf file.

```{r good-txt-1}
iris |>
  as_tibble() |>
  ggplot(aes(y = Petal.Width, x = Species)) + 
  geom_boxplot() +
  labs(title = "The title is super long and exceeds the word limits.
       Please do not take my points off.")

iris |>
  as_tibble() |>
  ggplot(aes(y = Petal.Width, x = Species)) + 
  geom_boxplot() +
  labs(title = "The title is super long and exceeds the word limits.",
       subtitle = "Please do not take my points off.")
```

Bad examples: text overflow when the functions have a lot of arguments needed specification

```{r bad-txt-2}
iris_boxplot <- iris |>
  as_tibble() |>
  ggplot(aes(y = Petal.Width, x = Species, fill = Species)) + 
  geom_boxplot() +
  labs(title = "Boxplot for petal width versus species", x = "species", y = "petal width", fill = "species")
```

A solution: represent them in separate lines

```{r good-txt-2}
iris_boxplot <- iris |>
  as_tibble() |>
  ggplot(aes(y = Petal.Width, x = Species, fill = Species)) + 
  geom_boxplot() +
  labs(title = "Boxplot for petal width versus species", 
       x = "species", 
       y = "petal width", 
       fill = "species")
```

#### 4. Inappropriate use of spacing

Bad example: +/-/=/\</\| are not surrounded by spaces

```{r bad-spacing-1}
iris_boxplot<-iris|>
  as_tibble()|>
  ggplot(aes(y=Petal.Width, x=Species, fill=Species))+ 
  geom_boxplot() 
```

Good example:

```{r good-spacing-1}
iris_boxplot <- iris |>
  as_tibble() |>
  ggplot(aes(y = Petal.Width, x = Species, fill = Species)) + 
  geom_boxplot() 
```

#### 5. New layer/pipline not begin with a newline

The code should be in a newline if it is a new function, such as pipeline, ggplot layers, model fitting.

Bad examples:

```{r bad-newline-1}
iris_boxplot <- iris |> as_tibble() |>
  ggplot(aes(y = Petal.Width, x = Species)) + 
  geom_boxplot() 
```

```{r bad-newline-2}
iris_boxplot <- iris |> 
  as_tibble() |>
  ggplot(aes(y = Petal.Width, x = Species)) + geom_boxplot() 
```

```{r bad-newline-3}
iris_boxplot <- iris |> 
  as_tibble() |>
  ggplot(aes(y = Petal.Width, x = Species)) + 
  geom_boxplot() + ggtitle("Boxplot for width versus species")
```

```{r bad-newline-4}
iris_sum <- iris |>
  as_tibble() |>
  lm(formula = Petal.Width ~ Sepal.Width) |> summary()
```

Good examples

```{r good-newline-3}
iris_boxplot <- iris |> 
  as_tibble() |>
  ggplot(aes(y = Petal.Width, x = Species)) + 
  geom_boxplot() + 
  ggtitle("Boxplot for width versus species")
```
