---
title: "Lab 5 - Logistic regression"
categories: "Lab"
execute:
  eval: true
editor: 
  markdown: 
    wrap: 80
---

::: callout-important
This lab is due Thursday, June 15th at 11:59pm.
:::

# Closing an Issue

Go to your GitHub repository. You should see an issue with the title "Learn to
close an issue with a commit". Your goal is to close this issue with a commit to
practice this workflow, which is the workflow you will use when you are
addressing feedback on your projects.

-   Go to Exercise 1 Part a in your lab .qmd file.
-   Delete the sentence that says "Delete this text!!!".
-   Render the document.
-   Commit your changes from the git tab with the commit message "Delete
    sentence, closes #1."
-   Push your changes to your repo and observe that the issue is now closed and
    the commit associated with this move is linked from the issue.

GitHub allows you to close an issue directly with commits if the commit uses one
of the following keywords followed by the issue number (which you can find next
to the issue title): close, closes, closed, fix, fixes, fixed, resolve,
resolves, and resolved.

# Lab

## Packages

You'll need the following packages for today's lab.

```{r}
#| warning: false
#| message: false
library(tidyverse)
library(tidymodels)
library(dsbox)
```

## Data

The data can be found in the **dsbox** package, and it's called `gss16`. Since
the dataset is distributed with the package, we don't need to load it
separately; it becomes available to us when we load the package.

If you would like to explicitly load the data into your environment so you can
view it, you can do so by running this code.

```{r}
#| warning: false
#| message: false

gss16 <- gss16
```

You can find out more about the dataset by inspecting its documentation, which
you can access by running `?gss16` in the Console or using the Help menu in
RStudio to search for `gss16`. You can also find this information
[here](https://rstudio-education.github.io/dsbox/reference/gss16.html).

# Exercises

## Exercise 1 - Data wrangling

::: callout-important
Remember: For each exercise, you should choose one person to type. All others
should contribute to the discussion, but only one person should type up the
answer, render the document, commit, and push to GitHub. All others should
**not** touch the document.
:::

a)  Create a new data frame called `gss16_advfront` that includes the variables
    `advfront`, `educ`, `polviews`, and `wrkstat`. Then, use the `drop_na()`
    function to remove rows that contain `NA`s from this new data frame. Sample
    code is provided below.

```{r}
#| eval: false
gss16_advfront <- gss16 |>
  select(___, ___, ___, ___) |>
  drop_na()
```

b)  Re-level the `advfront` variable such that it has two levels:
    `"Strongly agree"` and `"Agree"` combined into a new level called `"Agree"`
    and the remaining levels combined into `"Not agree"`. Then, re-order the
    levels in the following order: `"Agree"` and `"Not agree"`. Finally,
    `count()` how many times each new level appears in the `advfront` variable.

Hint: You can do this in various ways, but you'll likely need to use `mutate`
along with either `if_else()` or `case_when()` to re-level the variable and then
`fct_relevel()` to re-order the levels. (See Lab 2 for an example of using
`if_else` and HW 1 Exercise 4 for an example of using `fct_relevel`.)

c)  Combine the levels of the `polviews` variable such that levels that have the
    word "liberal" in them are lumped into a level called `"Liberal"` and those
    that have the word "conservative" in them are lumped into a level called
    `"Conservative"`. Then, re-order the levels in the following order:
    `"Conservative"` , `"Moderate"`, and `"Liberal"`. Finally, `count()` how
    many times each new level appears in the `polviews` variable.

::: callout-important
After the team member working on Exercise 1 renders, commits, and pushes, all
other team members should pull. Then, choose a new team member to write the
answer to Exercise 2. (And so on for the remaining exercises.)
:::

## Exercise 2 - Train and test sets

Now, let's split the data into training and test sets so that we can evaluate
the models we're going to fit by how well they predict outcomes on data that
wasn't used to fit the models.

Specify a random seed of 1234 (i.e., include `set.seed(1234)` at the beginning
of your code chunk), and then split `gss16_advfront` randomly into a training
set `train_data` and a test set `test_data`. Do this so that the training set
contains 80% of the rows of the original data.

## Exercise 3 - Logistic Regression

a)  Using the training data, specify a logistic regression model that predicts
    `advfront` by `educ`. In particular, the model should predict the
    probability that `advfront` has value `"Not agree"`. Name this model
    `model1`. Report the tidy output below.

b)  Write out the estimated model in proper notation. State the meaning of any
    variables in the context of the data.

c)  Using your estimated model, predict the probability of agreeing with the
    following statement: Even if it brings no immediate benefits, scientific
    research that advances the frontiers of knowledge is necessary and should be
    supported by the federal government (`Agree` in advfront) if you have an
    education of 7 years.

## Exercise 4 - Another model

a)  Again using the training data, fit a new model that adds the additional
    explanatory variable of `polviews`. Name this model `model2`. Report the
    tidy output below.

b)  Now, predict the probability of agreeing with the following statement: Even
    if it brings no immediate benefits, scientific research that advances the
    frontiers of knowledge is necessary and should be supported by the federal
    government (`Agree` in advfront) if you have an education of 7 years and are
    Conservative.

## Exercise 5 - Evaluating models with AIC

a)  Report the AIC values for each of `model1` and `model2`.

b)  Based on your results in part a, does it appear that including political
    views in addition to years of education is useful for modeling whether
    employees agree with the statement "Even if it brings no immediate benefits,
    scientific research that advances the frontiers of knowledge is necessary
    and should be supported by the federal government"? Explain.

## Exercise 6 - Evaluating models using test data

a)  For each of `model1` and `model2`, plot an ROC curve to visualize the true
    positive rate (sensitivity) and false positive rate (1 - specificity) for
    predictions on the observations in `test_data`. Add a title to each graph
    indicating which model they represent.

    You can use the code below to make the plot for model 1. Then adapt it as
    needed for model 2.

```{r}
#| eval: false

model1_pred <- predict(model1, test_data, type = "prob") |>  
  bind_cols(test_data |> select(advfront))

model1_pred |>
  roc_curve(
    truth = advfront, # advfront contains the true labels
    `.pred_Not agree`, # we're predicting the probability of "Not agree"
    event_level = "second" # "Not agree" is the 2nd level of advfront
  ) |>
  autoplot() +
  labs(title = "ROC curve for model 1")
```

b)  Report the area under the curve (AUC) for each of the ROC curves from part
    a.  
c)  Are the AUC values consistent with your conclusion in Exercise 5b? Explain.

# Submission

::: callout-warning
Before you wrap up the assignment, make sure all documents are updated on your
GitHub repo. We will be checking these to make sure you have been practicing how
to render and push changes.

You must turn in a PDF file to the Gradescope page by the submission deadline to
be considered "on time". Only one team member should submit to Gradescope, but
they should add all other team members to the submission.

Make sure your data are tidy! That is, your code should not be running off the
pages and spaced properly. See: <https://style.tidyverse.org/ggplot2.html>.
:::

To submit your assignment:

-   Go to <http://www.gradescope.com> and click *Log in* in the top right
    corner.
-   Click *School Credentials* $\rightarrow$ *Duke NetID* and log in using your
    NetID credentials.
-   Click on your *STA 199* course.
-   Click on the assignment, and you'll be prompted to submit it.
-   Mark all the pages associated with exercise. All the pages of your lab
    should be associated with at least one question (i.e., should be "checked").
    *If you do not do this, you will be subject to lose points on the
    assignment.*
-   Do not select any pages of your .pdf submission to be associated with the
    *"Workflow & formatting"* question.

# Grading

| Component             | Points |
|:----------------------|:-------|
| Ex 1                  | 10     |
| Ex 2                  | 10     |
| Ex 3                  | 5      |
| Ex 4                  | 10     |
| Ex 5                  | 5      |
| Ex 6                  | 5      |
| Workflow & formatting | 5      |
| **Total**             | 50     |

::: callout-note
The "Workflow & formatting" grade is to assess the reproducible workflow. This
includes:

-   linking all pages appropriately on Gradescope

-   putting your team and member names in the YAML at the top of the document

-   committing the submitted version of your `.qmd` to GitHub

-   Are you under the 80 character code limit? (You shouldn't have to scroll to
    see all your code). Pipes `%>%`, `|>` and ggplot layers `+` should be
    followed by a new line

-   You should be consistent with stylistic choices, e.g. only use 1 of `=` vs
    `<-` and `%>%` vs `|>`

-   All binary operators should be surrounded by space. For example `x + y` is
    appropriate. `x+y` is not.
:::
